---
title: "Uber Data Analysis"
author: "Seoyeong Park"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
#install.packages("ggthemes")
#install.packages("lubridate")
#install.packages("DT")
#install.packages("scales")
```

```{r}
library(dplyr)
```


## Uber Data Analysis with R
### Summary
This project is a visualization project to analyze trips by hours in a day and trips during months in a year. This project uses ggplot2 library and Uber dataset in NYC from Apr 2014 to Sep 2014.

### Importing necessary packages
```{r}
library(ggplot2)
library(ggthemes)
library(lubridate)
library(dplyr)
library(tidyr)
library(DT)
library(scales)
```

### Creating vector of colors to be implemented in plots
I will create a vector of colors which will be included in plotting functions. 
```{r}
colors = c("#2A3142", "#82CAC4", "#EE8A73", "#5B45AD", "#373728", "#8DAC60", "#C64300")
```

### Reading the data into designated variables
Now, I will read several csv files that contain the data prom April 2014 to September 2014. The data will be stored in corresponding data frames. After reading files, all data will be combined into a single dataframe.
Data is downloaded from below:
https://data-flair.training/blogs/r-data-science-project-uber-data-analysis/
```{r}
apr_data <- read.csv("uber-raw-data-apr14.csv")
may_data <- read.csv("uber-raw-data-may14.csv")
jun_data <- read.csv("uber-raw-data-jun14.csv")
jul_data <- read.csv("uber-raw-data-jul14.csv")
aug_data <- read.csv("uber-raw-data-aug14.csv")
sep_data <- read.csv("uber-raw-data-sep14.csv")

data_2014 <- rbind(apr_data, may_data, jun_data, jul_data, aug_data, sep_data)
```

```{r}
data_2014$Date.Time <- as.POSIXct(data_2014$Date.Time, format = "%m/%d/%Y %H:%M:%S")

data_2014$Time <- format(as.POSIXct(data_2014$Date.Time, format = "%m/%d/%Y %H:%M:%S"), format="%H:%M:%S")

data_2014$Date.Time <- ymd_hms(data_2014$Date.Time)
```

```{r}
data_2014$day <- factor(day(data_2014$Date.Time))
data_2014$month <- factor(month(data_2014$Date.Time, label = TRUE))
data_2014$year <- factor(year(data_2014$Date.Time))
data_2014$dayofweek <- factor(wday(data_2014$Date.Time, label=TRUE))
data_2014$hour <- factor(hour(hms(data_2014$Time)))
data_2014$minute <- factor(minute(hms(data_2014$Time)))
data_2014$second <- factor(second(hms(data_2014$Time)))
```

### Plotting the trips by the hours in a day
Using ggplot, I will plot the number of trips that passengers had made in a day. Through visualization, we can observe how the number of passengers fares in a day. 
```{r}
hour_data <- data_2014 %>%
  group_by(hour) %>%
  dplyr::summarize(Total = n())
datatable(hour_data)
```

```{r}
ggplot(hour_data, aes(hour, Total)) +
  geom_bar(stat = "identity", fill = "steelblue", color = "red") + 
  ggtitle("Trips Every Hour") +
  theme(legend.position = "none") +
  scale_y_continuous(labels = comma)

month_hour <- data_2014 %>%
  group_by(month, hour) %>%
  dplyr::summarize(Total = n())

ggplot(month_hour, aes(hour, Total, fill=month)) + 
  geom_bar(stat="identity") +
  ggtitle("Trips by Hour and Month") +
  scale_y_continuous(labels=comma)
```

### Plotting data by trips during every day of the month
```{r}
day_group <- data_2014 %>%
  group_by(day) %>%
  dplyr::summarize(Total = n())
datatable(day_group)

ggplot(day_group, aes(day, Total)) + 
  geom_bar(stat="identity", fill="lightblue") +
  ggtitle("Trips Every Day") + 
  theme(legend.position = "none") +
  scale_y_continuous(labels=comma)

day_month_group <- data_2014 %>%
  group_by(month, day) %>%
  dplyr::summarize(Total = n())

ggplot(day_month_group, aes(day, Total, fill=month)) +
  geom_bar(stat = "identity") +
  ggtitle("Trips by Day and Month") +
  scale_y_continuous(labels = comma) +
  scale_fill_manual(values=colors)
```

### Number of trips taking place during months in a year
```{r}
month_group <- data_2014 %>%
  group_by(month) %>%
  dplyr::summarize(Total =  n())
datatable(month_group)

ggplot(month_group, aes(month, Total, fill=month)) +
  geom_bar(stat="identity") +
  ggtitle("Trips by Month") +
  theme(legend.position = "none") +
  scale_y_continuous(labels = comma)+
  scale_fill_manual(values=colors)

month_weekday <- data_2014 %>%
  group_by(month, dayofweek) %>%
  dplyr::summarize(Total = n())

ggplot(month_weekday, aes(month, Total, fill=dayofweek)) +
  geom_bar(stat="identity", position="dodge")+
  ggtitle("Trips by Day and Month") +
  scale_y_continuous(labels = comma) + 
  scale_fill_manual(values = colors)
```

### Finding out the number of trips by bases
Visualize the number of trips that have been taken by passengers from each of the bases.
```{r}
ggplot(data_2014, aes(Base)) +
  geom_bar(fill="darkred") + 
  scale_y_continuous(labels=comma)+
  ggtitle("Trips by Bases")

ggplot(data_2014, aes(Base, fill=month)) +
  geom_bar(position="dodge") +
  scale_y_continuous(labels = comma) +
  ggtitle("Trips by Bases and Month") +
  scale_fill_manual(values = colors)

ggplot(data_2014, aes(Base, fill=dayofweek)) +
  geom_bar(position = "dodge") +
  scale_y_continuous(labels = comma) +
  ggtitle("Trips by Bases and DayofWeek") +
  scale_fill_manual(values=colors)
```


### Creating a heatmap of day, hour, month
I will create heatmaps using ggplot().
1. hour and day
2. month and day
3. month and day of the week
4. delinates month and bases
5. bases and day of the week
```{r}
day_and_hour <- data_2014 %>%
  group_by(day, hour) %>%
  dplyr::summarize(Total = n())

datatable(day_and_hour)

ggplot(day_and_hour, aes(day, hour, fill=Total)) +
  geom_tile(color = "white") +
  ggtitle("Heatmap by hour and day")

ggplot(day_month_group, aes(day, month, fill=Total)) +
  geom_tile(color = "white") +
  ggtitle("Heatmap by month and day")

ggplot(month_weekday, aes(dayofweek, month, fill=Total)) +
  geom_tile(color = "white") +
  ggtitle("Heatmap by month and day of week")

month_base <- data_2014 %>%
  group_by(Base, month) %>%
  dplyr::summarize(Total = n())

dayofweek_bases <- data_2014 %>%
  group_by(Base, dayofweek) %>%
  dplyr::summarize(Total = n())

ggplot(month_base, aes(Base, month, fill=Total)) +
  geom_tile(color = "white") +
  ggtitle("Heatmap by month and bases")

ggplot(dayofweek_bases, aes(Base, dayofweek, fill=Total)) +
  geom_tile(color="white") +
  ggtitle("Heatmap by bases and day of week")

```
